---
title: "Week 14 Discussion"
author: "Abby Kaff"
date: "12/12/2020"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(devtools)
install_github("vqv/ggbiplot")
library(ggbiplot)
```

We're going to do something different this week. Use the wholesale data and try to do some exploratory data analysis using some of the methods we've discussed (and other methods if you like). Focus on the quantitative variables, however, go ahead and have a look at how your clustering relates to the Channel data (you can ignore the Region variable).

Read in the data
```{r}
data <- read.csv("C:/Users/AK064060/Downloads/Wholesale customers data.csv")
# Split data into test/train
data$Channel <- as.factor(data$Channel)
divisionSize <- nrow(data) * 0.6
na.omit(data)
dataDivide <- sample(1:nrow(data),size=divisionSize)
train <- data[dataDivide,]
test<- data[-dataDivide,]
```

Run PCA
```{r}
pca_model <- prcomp(data[,2:8])
summary(pca_model)
```

```{r}
pca_model$center
pca_model$scale
pca_model$rotation

ggbiplot(pca_model,ellipse=TRUE, groups =data$Channel, obs.scale = 1, var.scale = 1)
```

```{r}
ggbiplot(pca_model, ellipse=TRUE, obs.scale = 1, var.scale = 1, var.axes = FALSE, groups = data$Channel)
```

Scree plot
```{r}
#Proportion of variance
prop_var <- ((pca_model$sdev)^2)/sum((pca_model$sdev)^2)
plot(prop_var, xlab = "Principal COmponent", ylab="Proportion of Variance Explained", type="b")

```
