---
title: "STAT 840 Lesson 8"
author: "Abby Kaff"
date: "12/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Lesson 8: Multiplicative Models, Models with Categorical Predictors, and Other Extensions

###Required Problems : 8.4, 8.11, 8.14, 8.15, 8.15, 8.21, 8.29, 8.40

####8.4 
Refer to Muscle mass Problem 1.27. A second-order regression model (8.2) with independent normal error terms is expected to be appropriate.

a) Fit regression model (8.2). Plot the fitted regression function and the data. Does the quadratic regression function appear to be a good fit here? Find $R^2$.

```{r}
library(readr)
CH01PR27 <- read_table2("C:/Users/AK064060/Desktop/CH01PR27.txt", 
    col_names = FALSE) #depends on computer being used
#CH01PR27 <- read_table2("~/Desktop/MS App Stats/STAT 840/Lesson 8 /CH01PR27.txt", col_names = FALSE)
colnames(CH01PR27)<-c("mass","age")
data<-CH01PR27
```

```{r}
# Fit the initial model in order to refine it
x1 = data$age-mean(data$age)
x1sq= x1^2
Data = cbind(data,x1,x1sq)

m1 <- lm(mass~x1+x1sq, data=Data)
summary(m1)
```
$y_i = 82.93-1.84_xi+ 0.0148_x_i^2$

b) Test whether or not there is a regression relation; use $\alpha = .05$. State the alternatives, decision rule, and conclusion. 

```{r}
anova(m1);qf(0.95,1,58)
```
The regression fuction looks to be a good fit, we should note the $R^2$ is small. Lack of fit F-stat is 0.62, p-value=0.757, the p-value is small, we fail to reject the null hypothesis (the relationship is reasonable). We do not have sufficient evidence to conclude that there is a lack of fit for the quadratic model. 


c) Estimate the mean muscle mass for women aged 48 years; use a 95 percent prediction interval. Interpret your interval. 

```{r}
a <- 48
s <- 2
n <- 20
error <- qnorm(0.975)*s/sqrt(n)
left <- a - error
right <- a + error
left;right
```
92.53 and 106.2

d) Predict the muscle mass for a woman whose age is 48 years; use a 95 percent prediction interval. Interpret your interval. 

```{r}
a <- 48
s <- 2
n <- 20
error <- qnorm(0.975)*s/sqrt(n)
left <- a - error
right <- a + error
left;right
```
80.0971 and 118.61

e) Test whether the quadratic term can be dropped from the regression model; use $\alpha = .05$. State the alternatives, decision rule, and conlusion. 

```{r}
t.star <- 1.776
t.star > 0.05 #do not reject H_0?
t.star < 2.00247 #do not reject H0?
```
We fail to reject the null hypothesis, so we do not need to drop the quadratic terms

f) Express the fitted regression function obtained in part (a) in terms of the original variable X. 

```{r}
#y_i = 82.93-1.84_xi+ 0.0148_x_i^2
```

g) Calculate the coefficient of simple correlation between $X$ and $X^2$ and between $x$ and $x^2$. Is the use of a centered variable helpful here?

The correlation between X and $X^2$ =.996 and the correlation between x and $x^2$ =-.0384. It is helpful because of the high correlation found in X and $X^2$.

####8.11 
Refer to Brand preference Problem 6.5. 

a) Fit regression model (8.22).

```{r}

```

b) Test whether or not the interaction term can be dropped from the model; use $\alpha = .05$. State the alternatives, decision rule, and conclusion. 

```{r}
```

####8.14 
In a regression study of factors affecting learning time for a certain task (measured in minutes), gender of learner was included as a predictor variable $(X_2)$ that was coded $X_2 = 1$ if male and 0 if female. It was found that $b_2 = 22.3$ and $s({b_2})=3.8$. An observer questioned whether the coding scheme for gender is fair becasue it results in a positive coefficient, leading to longer learning times for males than females. Comment. 

Yes I think in this instance it is considered a negative way to look at the coefficient, I would suggest changing it to a dummy variable of sex, one being male and one being female, in order to get a less skewed learning curve. 

####8.15
Refer to Copier Mintenance Problem 1.20. The users of the copiers are either training instituations that use a small model, or business firms that use a large, commercial model. An analyst at Tri-City wishes to fit a regression model including both number of copiers serviced $(X_1)$ and type of copier $(X_2)$ as predictor variables and estimate the effect of copier model (S--small, L--large) on number of minutes spent on the service call. Records show that the models serviced in the 45 calls were: 

Assume that regression model (8.33) is appropriate, and let $X_2 = 1$ if small model and 0 if large, commercial model. 

a) Explain the meaning of all regression coefficients in the model. 

y = response variable
$X_1$ = Predictor variable 1
$X_2$ = Predictor variable 2
e = residual error (unmeasured)
$\beta_0$ = Y intercept
$\beta_1$ = Regression coefficient 1
$\beta_2$ = Regression coefficient 2


b) Fit the regression model and state the estimated regression function. 

```{r}
data2 <- read_table2("C:/Users/AK064060/Desktop/CH01PR20.txt", 
    col_names = FALSE)
colnames(data2)<-c("X","Y")
X1 = data2$Y-mean(data2$Y)
X2 = X1^2
m2 <-lm(Y~X1+X2, data=data2)
summary(m2)
```
-0.5802+15.035(copiers serviced)+

c) Estimate the effect of copier model on mean service time with a 95 percent confidence interval. Interpret your interval estimate. 

```{r}
confint(m2,level=0.95)
```

d) Why would the analyst wish to include $X_1$, number of copiers, in the regression model when interest is in estimating the effect of type of copier model on service time? 



e) Obtain the residuals and plot them against $X_1X_2$.  IS there any indication that na interation term in the regression model would be helpful?

```{r}
plot(resid(m2))
```

####8.21
IN a regression analysis of on-the-job head injuries of warehouse laborers cause by falling objects, Y is a measure of severity of the injury, $X_1$ is an index reflecting both the weight of the object and the distance it fell, and $X_2$ and $X_3$ are indicator variables for nature of head protetion worn at the time of the accident:

The response function to be use in the study is \[E(Y) = \beta_0 + \beta_1X_1 + \beta_2X_2 + \beta_3X_3\].

a) Develop the response function for each type of protection category.

```{r}
```

b) For each o the following questions, specify the alternatives $H_0$ and $H_a$ for the appropriate test: (1) With $X_1$ fixed, does wearing a bump cap reduce the expected severity of injury as compared with wearing no protection? (2) With $X_1$ fixed, is the expected severity of injury the same when wearing a hard hat as when wearing a bump cap?

```{r}
```

####8.29
Consider the second-order regession model with one predictor variable in (8.2) and the following two sets of X values: 

Set 1: 1.0  1.5  1.1  1.3   1.9  .8  1.2  1.4  
Set 2: 12   1   123   17    415   71  283   38

For each set, calculate the coefficient of correlation between $X$ and $X^2$, then between $x$ and $x^2$. Also calculate the coefficients of correlation between $X$ and $X^3$ and between $x$ and $x^3$. What generalizations are suggested by your results? 

```{r}
```
